Tulai

### 2023-09-29 paper/paper.qmd
add some chi-square tests and residual plots

```{r}
# load libraries
library(tidyverse)
library(here)
library(ggpmisc)
library(ggrepel)
library(ggbeeswarm)
library(cowplot)
library(knitr)
```

@@

@@

@@

kable(
  summary_table_tool_area,
  col.names = c("Tools", "D1", "TA", "TP1", "Total"),
  caption = "Tool Typology in the three main areas")

@@

>>>>

```{r}
# load libraries
library(tidyverse)
library(here)
library(ggpmisc)
library(ggrepel)
library(ggbeeswarm)
library(cowplot)
library(knitr)
library(ggcorrplot)
library(ggpubr)
library(ggbiplot)
```

@@

# chi-square test for different frequencies of tool type groups 
# by area
  summary_table_area_filtered_for_chi_square <- 
  summary_table_area_filtered %>% 
  pivot_longer(-area) %>% 
  pivot_wider(names_from = "area") %>% 
  select(-name)

chisq <- 
  chisq.test(summary_table_area_filtered_for_chi_square)

rownames(chisq$residuals) <- 
   summary_table_area_filtered %>% 
  pivot_longer(-area) %>% 
  pivot_wider(names_from = "area") %>% 
  pull(name)  %>% 
  str_remove("\\(n\\)")

ggcorrplot(chisq$residuals,
           method = 'circle') + 
  ggplot2::scale_fill_gradient2(
    low = 'red',
    high = 'royalblue',
    mid = 'white',
    midpoint = 0,
    limit = c(-7, 11),
    space = "Lab",
    name = "Residual"
  ) + 
  coord_flip()

@@

A Chi-square test on the frequency of major lithic types by excavation area reveals a significant difference between the areas (`r apastats::describe.chi(summary_table_area_filtered_for_chi_square)`). Figure XXX shows that there are fewer utlized pieces and flakes in d1 than expected, and more flakes and fewer cores and blades in tp1 than expected assuming a null hypothesis of no difference between the areas.

@@

kable(
  summary_table_tool_area,
  col.names = c("Tools", "D1", "TA", "TP1", "Total"),
  caption = "Tool Typology in the three main areas")
```

# chi-square test for different frequencies of tool type groups 
# by area
  summary_table_tool_area_for_chi_square <- 
  summary_table_tool_area %>% 
  filter(!tooltypegroup == "Total") %>% 
  select(d1, ta, tp1)

chisq <- 
  chisq.test(summary_table_tool_area_for_chi_square)

rownames(chisq$residuals) <- 
  summary_table_tool_area %>% 
  filter(!tooltypegroup == "Total") %>% 
  pull(tooltypegroup)

ggcorrplot(chisq$residuals,
           method = 'circle') + 
  ggplot2::scale_fill_gradient2(
    low = 'red',
    high = 'royalblue',
    mid = 'white',
    midpoint = 0,
    limit = c(-7, 11),
    space = "Lab",
    name = "Residual"
  ) + 
  coord_flip()

@@

A Chi-square test on the frequency of tool type groups by excavation area reveals a significant difference between the areas (`r apastats::describe.chi(summary_table_tool_area_for_chi_square)`). Figure XXX shows that there are more geomeric pieces and burins in d1 than expected, and more backed and retouched pieces in tp1 than expected assuming a null hypothesis of no difference between the areas.
